Structured Versus Decorated Cospans

### 2021-04-19 cospans_new.tex
edited Kenny's proof of Theorem 2.1 up to the material in red

In what follows, we define a \define{cocartesian monoidal category} $(\A,+)$ to be a symmetric
monoidal category where the monoidal structure is given by chosen binary coproducts, with a chosen initial object serving as the unit for the monoidal structure.  Any category with finite coproducts can be made into a cocartesian monoidal category by arbitrarily choosing finite coproduct and a unit object.

%In all that follows, when we say a category `has finite colimits' we mean it is equipped with a choice of colimit for every finite diagram.   Thus, if $\A$ has coproducts it gives a cocartesian monoidal category $(\A,+)$: that is, a symmetric monoidal category where the monoidal structure is given by the chosen binary coproducts and initial object.   However, when we say a functor `preserves finite colimits', it need only do this up to canonical isomorphism, unless otherwise specified.   MAY NEED THIS LATER, BUT NOT HERE!!!

\begin{thm}\label{thm:decorated_cospans_1}
Let $(\A,+)$ be cocartesian monoidal and $F \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\begin{equation}\label{eq:mapofdec}
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

We prove both these theorems using the work of Shulman \cite{Shulman2008}, which we recall here.   First we need a few concepts.   Opfibrations are recalled in \cref{subsec:fibrations}, 
and fibrations are defined dually, with a functor $\Phi \maps \C \to \A$ being a fibration if and only
if $\Phi\op \maps \C\op \to \A\op$ is an opfibration.   A functor $\Phi \maps \C \to \A$ is a \define{bifibration} if it is both a fibration and an opfibration.      If $\C$ and $\A$ are monoidal categories, a \define{monoidal} bifibration $\Phi \colon \C \to \A$ is a bifibration that is also a strict monoidal functor.   Similarly, if $\C$ and $\A$ are symmetric monoidal, a \define{symmetric monoidal} bifibration is a bifibration $\Phi \colon \C \to \A$ that is a symmetric strict monoidal functor.

If $\Phi \maps \C \to \A$ an opfibration, for any morphism $f \maps a \to b$ in $\A$ there is a covariant reindexing functor
\[    f_! \maps \C_a \to \C_b \]
between the fibers of $\Phi$, as recalled in \cref{subsec:fibrations}.  Similarly, if $\Phi$ is a fibration there is a contravariant reindexing functor
\[    f^\ast \maps \C_b \to \C_a .\]
Thus, in a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f^\ast$ is left adjoint to $f_\ast$ \cite[Proposition 3.9]{Shulman2008}.   Using this one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a specified natural transformation, built
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\C_a$};
\node (B) at (1.5,0) {$\C_b$};
\node (C) at (0,-1) {$\C_c$};
\node (D) at (1.5,-1) {$\C_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
If $\theta$ is a natural isomorphism whenever the original square in $\A$ is
a pushout, we say $\Phi$ is \define{Beck--Chevalley}.  (Shulman says `strongly co-Beck--Chevalley'.)

@@

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
We apply \cref{lem:shulman} to a bifibration $\Phi \maps (\A/U) \to \A$ which we now define. The functor $U \colon \inta F \to \A$ is the associated monoidal opfibration of the symmetric lax monoidal pseudofunctor $F \colon \A \to \Cat$. Let $(\A/U)$ be the comma category in which:
\begin{enumerate}
\item{an object is a triple $(a,x,a \xrightarrow{i} U(x))$ with $a \in \A$ and $x=(a',p) \in \A \times F(a')$, and}
\item{a morphism from $(a,x,a \xrightarrow{i} U(x))$ to $(b,x',b \xrightarrow{i'} U(x'))$ is a pair of morphisms $f \colon a \to b$ and $\alpha \colon x \to x'$ such that the following square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$U(\alpha)$} (D);
\end{tikzpicture}
\]}
\end{enumerate}
The functor $\Phi$ is defined on objects by $\Phi(a,x,a \xrightarrow{i} U(x))=a$ and on morphisms by $\Phi(f,\alpha)=f$. We have a monoidal structure on the comma category $(\A/U)$ where the tensor product of two objects $(a_1,x_1,a_1 \xrightarrow{i_1} U(x_1))$ and $(a_2,x_2,a_2 \xrightarrow{i_2} U(x_2))$ is given by
\[  (a_1,x_1,a_1 \xrightarrow{i_1} U(x_1)) \otimes (a_2,x_2,a_2 \xrightarrow{i_2} U(x_2)) = (a_1+a_2,x_1 \otimes x_2,a_1+a_2 \xrightarrow{i_1+i_2} U(x_1 \otimes x_2)) \]
where we have left the equality $U(x_1)+U(x_2) = U(x_1 \otimes x_2)$ implicit in the above composite, for reasons of space. The monoidal unit is given by $(0_\A,0_\X,0_\A \to U(0_\X))$. It is easy to see that the functor $\Phi$ can be made into a strict monoidal functor using this tensor product.

The functor $\Phi$ is a fibration where given a morphism $f \maps a \to b$ in $\A$ with $(b,y,b \xrightarrow{i} U(y))$ over $b$ in $\int F$, the cartesian lifting of $f$ is given by 
\[ (f,\id) \maps (a,y,a \xrightarrow{i \circ f} U(y)) \to (b,y,b \xrightarrow{i} U(y)). \]
Denoting the cartesian lifting of a morphism $f$ by $\textnormal{cart}(f)$, the tensor product of $(\A/U)$ strictly preserves cartesian liftings in the sense that $\textnormal{cart}(f_1) \otimes \textnormal{cart}(f_2)=\textnormal{cart}(f_1+f_2)$.

The functor $\Phi$ is also an opfibration where given a morphism $f \maps a \to b$ with $(a,x,a \xrightarrow{i} U(x))$ over $a$, the cocartesian lifting of $f$ is given by 
\[ (f,\theta) \maps (a,x,a \to U(x)) \to (b,z,b \xrightarrow{\phi} U(z)) \]
where $\theta$ and $\phi$ arise from the following pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x)+_a b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$\phi$} (D)
(C) edge [dashed] node[above]{$\theta$} (D);
\end{tikzpicture}
\]
and the object $z$ arises from the cocartesian lifting of $\theta$ through $U$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (2.5,0) {$c$};
\node (C) at (1.5,1) {$b$};
\node (D) at (0,-2) {$U(x)$};
\node (E) at (2.5,-2) {$U(y)$};
\node (F) at (1.5,-1) {$U(z) = U(x)+_a b$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$\theta$} (F)
(F) edge [dashed] node [left] {$\exists ! \psi$} (E)
(D) edge node [above] {$U(g')$} (E)
(A) edge node [left] {$i$} (D)
(C) edge [dashed] node [right,near end] {$\phi$} (F)
(B) edge node {$$} (E)
(A) edge node[above]{$f'$} (B)
(C) edge node [right] {$p$} (B)
(A) edge node[left]{$f$} (C);
%(B) edge [dashed] node[right]{$$} (D)
%(C) edge [dashed] node[above]{$$} (D);
\end{tikzpicture}
\]
Denoting the cocartesian lifting of a morphism $f$ by $\textnormal{cocart}(f)$, we have that $\textnormal{cocart}(f_1) \otimes \textnormal{cocart}(f_2) \cong \textnormal{cocart}(f_1+f_2)$ by the universal map between two colimits of the same diagram: 
\[  (U(x_1)+_{a_1}b_1)+(U(x_2)+_{a_2}b_2) \cong U(x_1 \otimes x_2)+_{a_1+a_2}(b_1+b_2). \]

Lastly, the monoidal bifibration $\Phi$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$(\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$ where $x=(b',p) \in \A \times F(b')$ so that $U(x)=b'$. Going left, we precompose with $h \colon a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \xrightarrow{\psi} U(z))$ of $(\A/U)_c$, where $z \in \A \times F(c+_a U(x))$ is given by $z=(c+_a U(x),q)$ where $q$ is given by
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_c +_{!_a} 1)} F(c+_a U(x)). \]
%and so by construction, we have that 
%\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
%{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \colon b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \A \times F(d+_b U(x))$ is given by ${z'=(d+_b U(x),q')}$ where $q'$ is given by 
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_d +_{!_b} 1)} F(d+_b U(x)). \]
We then precompose with $f \colon c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting: given a (symmetric) monoidal category $\A$, there is a bijection between \define{(symmetric) monoidal opindexed categories}, which are (symmetric) lax monoidal pseudofunctors $F\maps (\A, \otimes_A) \to (\Cat, \times)$, and \define{(symmetric) monoidal opfibrations}, which are opfibrations $U \maps \X \to \A$ where $\X$, $\A$ are (symmetric) monoidal, $U$ is (symmetric) strict monoidal, and $\otimes_\X$ preserves cocartesian liftings.   This bijection sends a monoidal opindexed category $F \maps (\A,\otimes_A) \to (\Cat, \times)$ to the monoidal opfibration $U \maps \inta F \to \A$.  If the monoidal structure on $\A$ is in fact cocartesian, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps \A \to \MonCat$ (or to $\SMC$ in the symmetric case).

In fact, all these correspondences lift to 2-equivalences between appropriate 2-categories.

>>>>

In all that follows, when we say a category `has finite colimits' we mean it is equipped with a choice of colimit for every finite diagram.   Thus, if $\A$ has coproducts it gives a cocartesian monoidal category $(\A,+)$: that is, a symmetric monoidal category where the monoidal structure is given by the chosen binary coproducts and initial object.   However, when we say a functor `preserves finite colimits', it need only do this up to canonical isomorphism, unless otherwise specified.

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $F \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\begin{equation}\label{eq:mapofdec}
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

We prove both these theorems using the work of Shulman \cite{Shulman2008}, which we recall here.   First we need a few concepts.   Opfibrations are recalled in \cref{subsec:fibrations}, and fibrations are defined dually, with a functor $\Phi \maps \C \to \A$ being a fibration if and only if $\Phi\op \maps \C\op \to \A\op$ is an opfibration.   A functor $\Phi \maps \C \to \A$ is a \define{bifibration} if it is both a fibration and an opfibration.      If $\C$ and $\A$ are monoidal categories, a \define{monoidal} bifibration $\Phi \maps \C \to \A$ is a bifibration that is also a strict monoidal functor.   Similarly, if $\C$ and $\A$ are symmetric monoidal, a \define{symmetric monoidal} bifibration is a bifibration $\Phi \maps \C \to \A$ that is a symmetric strict monoidal functor.

If $\Phi \maps \C \to \A$ an opfibration, for any morphism $f \maps a \to b$ in $\A$ there is a covariant reindexing functor
\[    f_! \maps \C_a \to \C_b \]
between the fibers of $\Phi$, as recalled in \cref{subsec:fibrations}.  Similarly, if $\Phi$ is a fibration there is a contravariant reindexing functor
\[    f^\ast \maps \C_b \to \C_a .\]
Thus, in a bifibration we have both covariant and contravariant reindexing functors, and in 
fact $f^\ast$ is left adjoint to $f_\ast$ \cite[Proposition 3.9]{Shulman2008}.   Using this one can easily show that for any commutative square in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
%\node (E) at (0.5,-0.5) {$\Downarrow\theta$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
the following square commutes up to a specified natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\C_a$};
\node (B) at (1.5,0) {$\C_b$};
\node (C) at (0,-1) {$\C_c$};
\node (D) at (1.5,-1) {$\C_d$};
\node (E) at (0.75,-0.5) {$\Searrow \theta$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^\ast$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^\ast$} (C);
\end{tikzpicture}
\]
where $\theta$ is built as a composite involving the unit of the adjunction between
$g_!$ and $g^\ast$ and the counit of the adjunction between $k_!$ and $k^\ast$:
\[    k_! h^\ast \To k_! h^\ast g^\ast g_! = k_! k^\ast f^\ast g^! \To f^\ast g_! .\] 
If $\theta$ is a natural isomorphism whenever the original square in $\A$ is
a pushout, we shall say $\Phi$ is \define{Beck--Chevalley}.  (Shulman says `strongly co-Beck--Chevalley'.)

@@

We  prove Theorems \ref{thm:decorated_cospans_1} and \ref{thm:decorated_cospans_2} by applying this lemma to a particular monoidal bifibration.  To get our hands on this starting from our lax monoidal pseudofunctor $F \maps (A,+) \to (\Cat,\times)$, we need a monoidal version of the Grothendieck construction \cite{MV}.   We recall the Grothendieck construction in \ref{subsec:fibrations}; briefly,
this gives a bijection between pseudofunctors $F \maps \A \to \Cat$ and opfibrations
$U \maps \inta F \to \A$.   Moeller and the third author \cite{MV} showed that
similarly, given a (symmetric) monoidal category $\A$, there is a bijection between 
\begin{itemize}
\item \define{(symmetric) monoidal opindexed categories}: (symmetric) lax monoidal pseudofunctors \hfill \break $F\maps (\A, \otimes_A) \to (\Cat, \times)$
\end{itemize}
and 
\begin{itemize}
\item \define{(symmetric) monoidal opfibrations}: opfibrations $U \maps \X \to \A$ where $\X$ and $\A$ are (symmetric) monoidal, $U$ is (symmetric) strict monoidal, and $\otimes_\X$ preserves cocartesian liftings. 
\end{itemize}  
If the monoidal structure on $\A$ is in fact cocartesian, as it is in our case, there is a further correspondence between these structures and ordinary pseudofunctors $F\maps \A \to \MonCat$ (or to $\SMC$ in the symmetric case).

In fact, all these correspondences lift to 2-equivalences between appropriate 2-categories:

\begin{lem}
\label{lem:MonGroth}
There is a 2-equivalence between the 2-categories of monoidal opfibrations and monoidal opindexed categories, and if the base is cocartesian monoidal, there is a 2-equivalence between these and pseudofunctors into $\MonCat$.   Similarly there is a 2-equivalence between symmetric monoidal opfibrations and symmetric monoidal opindexed categories, and if the base is cocartesian monoidal, also a 2-equivalence between these and pseudofunctors into $\SMC$.   
\end{lem}

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
To obtain the desired double category we apply \cref{lem:shulman} to a bifibration $\Phi \maps \A/U \to \A$ which we now define.   Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps \A \to \Cat$.   Thus:
\begin{itemize}
\item an object of $\inta F$ is a pair $(b,p)$ where $b \in \A$ and $p \in F(A)$;
\item a morphism in $\inta F$ from $(b,p)$ to $(b',p')$ is a pair $(g,\zeta)$
where $g \maps b \to b'$ is a morphism in $\A$ and $\zeta \maps F(f)(p) \to p'$ is
a morphism in $F(b')$.
\end{itemize}
The functor $U$ maps objects and morphisms to their first component.   We thus obtain an explicit description of the comma category $\A/U$:
\begin{itemize}
\item an object of $\A/U$ is a triple $(a,x,a \xrightarrow{i} U(x))$ with $a \in \A$ and $x=(b,p) \in \inta F$;
\item a morphism in $\A/U$ from $(a,x,a \xrightarrow{i} U(x))$ to $(a',x',a' \xrightarrow{i'} U(x'))$ is a pair $(f,\theta)$ where $f \maps a \to a'$ is a morphism in $\A$ and $\theta \maps x \to x'$ is a morphism in $\inta F$ such that the following square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (1.5,-1) {$U(x')$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$U(\theta)$} (D);
\end{tikzpicture}
\]
\end{itemize}
The functor $\Phi \maps \A/U \to \A$ is defined on objects by $\Phi(a,x,a \xrightarrow{i} U(x))=a$ and on morphisms by $\Phi(f,\alpha)=f$. We have a monoidal structure on $\A/U$ where the tensor product of two objects $(a_1,x_1,a_1 \xrightarrow{i_1} U(x_1))$ and $(a_2,x_2,a_2 \xrightarrow{i_2} U(x_2))$ is given by
\[  (a_1,x_1,a_1 \xrightarrow{i_1} U(x_1)) \otimes (a_2,x_2,a_2 \xrightarrow{i_2} U(x_2)) = (a_1+a_2,x_1 \otimes x_2,a_1+a_2 \xrightarrow{i_1+i_2} U(x_1 \otimes x_2)) \]
where we have left the equality $U(x_1)+U(x_2) = U(x_1 \otimes x_2)$ implicit in the above composite, for reasons of space. The monoidal unit is given by $(0_\A,0_\X,0_\A \to U(0_\X))$. It is easy to see that the functor $\Phi$ is a strict monoidal using this tensor product.

One can easily check that $\Phi$ is a fibration.  Given a morphism $f \maps a \to b$ in $\A$ and an object $(b,y,b \xrightarrow{i} U(y)) \in \A/U$ over $b$, a cartesian lifting of $f$ is given by 
\[ (f,\id) \maps (a,y,a \xrightarrow{i \circ f} U(y)) \, \to \, (b,y,b \xrightarrow{i} U(y)). \]
Denoting this cartesian lifting by $\textnormal{cart}(f)$, the tensor product of $\A/U$ strictly preserves such cartesian liftings in the sense that $\textnormal{cart}(f_1) \otimes \textnormal{cart}(f_2)=\textnormal{cart}(f_1+f_2)$.

One can also check that the functor $\Phi$ is also opfibration, but this takes a bit more work.  Given a morphism $f \maps a \to a'$ in $\A$ and an object $(a,x,a \xrightarrow{i} U(x)) \in \A/U$ over $a$, we shall show that a cocartesian lifting of $f$ is given by 
\[ (f,\theta) \maps (a,x,a \xrightarrow{i} U(x)) \, \to \, (a',x', a' \xrightarrow{i'} U(x')). \]
Here $i'$ arises from the following pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (2,0) {$a'$};
\node (C) at (0,-1) {$U(x)$};
\node (D) at (2,-1) {$U(x)+_a a'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge [dashed] node[right]{$i'$} (D)
(C) edge [dashed] node[below]{$g$} (D);
\end{tikzpicture}
\]
The object $x' \in \inta F$ is defined as follows.  Recall that $x \in \inta F$ is a
pair $(b,p)$ with $b\in \A$ and $p \in F(b)$, and note that $U(x) = b$.   Define 
$x'$ to be the pair $(b',p')$ where $b' = U(x)+_a a'$ is the pushout in the above
diagram and $p' = F(g)(p)$.  Then $U(x') = b'$ so we have 
$i' \maps a' \to U(x')$ as desired.  Finally, we define $\theta \maps x \to x'$ to be 
the morphism in $\inta F$ given by the pair $(g,\id)$ where $g \maps b \to b'$ is
given in the above diagram and $\id \maps F(f)(p) \to p'$ is the identity.

{\john To show that $(f,\theta)$ is a cocartesian lifting of $f$, suppose...
Then use this picture somehow:}
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (2.5,0) {$c$};
\node (C) at (1.5,1) {$a'$};
\node (D) at (0,-2) {$U(x)$};
\node (E) at (2.5,-2) {$U(y)$};
\node (F) at (1.5,-1) {$U(x') = U(x)+_a a'$};
\path[->,font=\scriptsize,>=angle 90]
(D) edge node [above] {$\theta$} (F)
(F) edge [dashed] node [right] {$\exists ! \psi$} (E)
(D) edge node [above] {$U(g')$} (E)
(A) edge node [left] {$i$} (D)
(C) edge node [right,near end] {$i'$} (F)
(B) edge node {$$} (E)
(A) edge node[above]{$f'$} (B)
(C) edge node [right] {$p$} (B)
(A) edge node[left]{$f$} (C);
%(B) edge [dashed] node[right]{$$} (D)
%(C) edge [dashed] node[above]{$$} (D);
\end{tikzpicture}
\]

Denoting the cocartesian lifting of a morphism $f$ by $\textnormal{cocart}(f)$, we have that $\textnormal{cocart}(f_1) \otimes \textnormal{cocart}(f_2) \cong \textnormal{cocart}(f_1+f_2)$ by the universal map between two colimits of the same diagram: 
\[  (U(x_1)+_{a_1}b_1)+(U(x_2)+_{a_2}b_2) \cong U(x_1 \otimes x_2)+_{a_1+a_2}(b_1+b_2). \]

Lastly, the monoidal bifibration $\Phi$ satisfies the Beck--Chevalley condition. Given a pushout square in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$b$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$d$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$g$} (D)
(C) edge node[below]{$f$} (D);
\end{tikzpicture}
\]
we need to show that the following square commutes up to a natural isomorphism:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(\A/U)_a$};
\node (B) at (1.5,0) {$\A/U)_b$};
\node (C) at (0,-1) {$(\A/U)_c$};
\node (D) at (1.5,-1) {$(\A/U)_d$};
\node (E) at (0.75,-0.5) {$\Searrow$};
\path[->,font=\scriptsize,>=angle 90]
(B) edge node[above]{$h^*$} (A)
(A) edge node[left]{$k_!$} (C)
(B) edge node[right]{$g_!$} (D)
(D) edge node[below]{$f^*$} (C);
\end{tikzpicture}
\]
Let $(b,x,b \xrightarrow{i} U(x))$ be an object of $(\A/U)_b$ where $x=(b',p) \in \A \times F(b')$ so that $U(x)=b'$. Going left, we precompose with $h \maps a \to b$ to obtain the object $(a,x,a \xrightarrow{i \circ h} U(x))$ of $(\A/U)_a$. Then going down, we push forward along $k$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$c$};
\node (D) at (1.5,-1) {$c+_{a}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i \circ h$} (B)
(A) edge node[left]{$k$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\psi$} (D);
\end{tikzpicture}
\]
to obtain the object $(c,z,c \xrightarrow{\psi} U(z))$ of $(\A/U)_c$, where $z \in \A \times F(c+_a U(x))$ is given by $z=(c+_a U(x),q)$ where $q$ is given by
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_c +_{!_a} 1)} F(c+_a U(x)). \]
%and so by construction, we have that 
%\[  U(z)=U(\bot_c+_{\bot_a}x)=c+_a U(x).\]
%{\john Need to refer to a proposition where we show $U$ preserves pushouts strictly.}
 For the other route, first going down, we push the object $(b,x,b \xrightarrow{i} U(x))$ forward along $g \maps b \to d$ by taking the following pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$U(x)$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}U(x)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\theta$} (D);
\end{tikzpicture}
\]
which yields the object $(d,z',d \xrightarrow{\theta} U(z'))$ of $(\A/U)_d$ where $z' \in \A \times F(d+_b U(x))$ is given by ${z'=(d+_b U(x),q')}$ where $q'$ is given by 
\[ 1 \xrightarrow{p} F(U(x)) \xrightarrow{F(\zeta)} F(0+_0 U(x)) \xrightarrow{F(!_d +_{!_b} 1)} F(d+_b U(x)). \]
We then precompose with $f \maps c \to d$ to obtain the object $(c,z',c \xrightarrow{\theta \circ f} U(z'))$ of $(\A/U)_c$.

@@

FIX ABOVE PARAGRAPH.

The 2-equivalence of \cref{lem:MonGroth} has an important special case where the Grothendieck category $(X,\otimes_\X)$ is cocartesian monoidal itself, with coproducts built up from \cref{eq:explicitstructure2}.  In that case, opfibrations $(\X,+) \to (\A,+)$ that strictly preserve coproducts and initial object bijectively correspond to pseudofunctors into the 2-category of cocartesian categories.  For more details, see \cite[Corollary 4.7]{MV} and the related discussion.

We can further restrict to the case of opfibrations that also preserve pushouts, and thus all finite colimits, thanks to the following more general result.
